X-ray phase-contrast imaging at \SI{100}{\kilo\electronvolt} on conventional sources

### 2013-10-28 hedpc_natphys.tex
Marcos revision for the paper

X-ray radiography and \ac{CT} are standard imaging techniques in
materials and life sciences for the nondestructive examination of samples or
the diagnosis of diseases in patients. The underlying contrast mechanism
relies on the different X-ray attenuation properties of different materials
or tissue types. The dominant physical effects contributing to attenuation
are the photoelectric effect and incoherent (Compton) scattering. Besides attenuation, the wave
nature of X-rays reveals another contrast mechanism, which is the phase
shift. The interaction contributing to phase shifts is coherent (Rayleigh)
scattering~\cite{Als-Nielsen2011}.

@@

Phase-sensitive
imaging is a desirable modality, as it can provide a complementary source of
contrast with respect to absorption by providing direct access 
to the electron density~\cite{Als-Nielsen2011}. Moreover, the combination with
attenuation enables the determination of the effective atomic
number~\cite{Qi2010}.
An enhanced
\ac{CNR} in images compared to attenuation for certain
materials or tissues~\cite{Pfeiffer2007a,McDonald2009} has also been
demonstrated.

The vast majority of phase-sensitive techniques, including crystal analyzer
based~\cite{Davis1995,Chapman1997} or interferometric~\cite{Bonse1965,Momose1996}
methods rely on X-ray beams of high spatial and temporal coherence, which is
available only at synchrotron sources. Inline
phase contrast~\cite{Snigirev1995,Wilkins1996,Cloetens1996} and
Talbot interferometry~\cite{Cloetens1997,David2002,Momose2003a} need high
spatial coherence but are available on polychromatic microfocus sources.
Phase-contrast imaging using X-ray beams of low temporal \emph{and}
spatial coherence such as conventional low-brilliance X-ray tubes have been
demonstrated with coded
apertures~\cite{Munro2012} and Talbot-Lau
interferometry~\cite{Pfeiffer2006}. Analyzer-based systems have been
recently extended to tube sources~\cite{Nesch2009,Parham2009}
but the photon energy is limited
to the tungsten $K_\alpha$ line at~\SI{60}{\kilo\eV}.
In addition to phase sensitivity, these techniques provide a dark-field
contrast, also called scatter contrast or visibility-reduction contrast,
that is proportional to the integrated local small angle scattering power
from microscopic density fluctuations in a specimen~\cite{Pfeiffer2008}.

Talbot interferometry has been demonstrated using a synchrotron source at
$\SI{82}{\kilo\electronvolt}$~\cite{Willner2013} and
$\SI{123}{\kilo\electronvolt}$~\cite{Ruiz2013}.
Using a low-brilliance X-ray tube, Talbot-Lau interferometry was
applied at $\SI{60}{\kilo\electronvolt}$ mean energy~\cite{Donath2009}.
Medical imaging applications may benefit from phase contrast at higher
energies: chest or abdominal
radiography or \ac{CT} require energies between $100$ and
$\SI{150}{\kilo\voltpeak}$. Other potential applications are homeland
security or chip failure analysis, which require high energies for the
visualization of materials of high density and atomic number.

Here, we introduce a method for phase-contrast imaging which is
compatible with the entire diagnostic energy range of X-rays, 
compact imaging arrangements and 
conventional X-ray tubes. The method is based on Talbot-Lau
interferometry~\cite{Pfeiffer2006} and employs an edge-on illumination approach for the grating design and
arrangement. 
The challenge which limited the progress towards higher energies was mainly
related to the manufacturing of gratings with high aspect ratios. The aspect
ratio, given by
\begin{equation}
    \text{R} = \frac{2h}{p},
\end{equation}
where $p$ is the grating period and $h$ the  structure
height, is normally limited by the lithographic process, as grating
structures tend to collapse or deform (e.g. due to capillary forces) if
the aspect ratio is too high. For a given setup length these parameters
depend on the target energy $E$ according to $p \propto
1/\sqrt{E}$ and $h \propto E^3$, and therefore $\textnormal{R}
\propto E^{7/2}$~\cite{Momose2003a}. If at $E=\SI{25}{\kilo\electronvolt}$ an aspect ratio
for the absorption grating around $\textnormal{R}=30$ is necessary for a
reasonable length of the experimental arrangement, it would have to be at
least $128$ for $E=\SI{100}{\kilo\electronvolt}$. Moreover, when using a
broad spectrum, photons above the design energy should also be
efficiently blocked by the gratings, which requires even higher aspect ratios.
The maximum achievable aspect ratios of current fabrication techniques~\cite{David2007,Kenntner2010} are around 60.

Our design introduces the edge-on
illumination of circularly aligned structures. Edge-on illumination
(figure~\ref{Fig:schematic}), as
opposed to face-on illumination, exploits the dimension along the grating
lines to form a high aspect ratio of the structures in the direction of the beam. The
effective structure height of the grating is then determined by the grating
dimension along the grating lines, which essentially allows arbitrarily high
aspect ratios. 
\begin{figure}[ht]
    \includegraphics[width=.7\textwidth]{figures/figure1.eps}
    \caption{Schematic of a grating
        interferometer for X-ray energies between 60 and
        \SI{150}{\kilo\electronvolt} in edge-on illumination mode. The
        aspect ratio is defined by the ratio of the travelling distance along the
        grating lines and the period and can be arbitrarily long. In order to avoid
        a reduction of the field of view, the grating structures are aligned on an
    arc.}
    \label{Fig:schematic}
\end{figure}

Increasing the aspect ratio of the gratings typically leads to a
reduction of the field of view due to the change of the grating transmission
function at high incident angles. In order to overcome this problem, the grating lines are circularly aligned 
with a radius equal to the distance to the source.

The combination of edge-on illumination and circularly aligned structures
enables phase-contrast imaging at arbitrary design energies and with a
maximum field of view in the horizontal direction ($x$ direction). These
advantages come at the expense of a limited field of view in the vertical
direction ($y$ direction), which is, depending on the X-ray detector,
typically a few pixels. However, radiographic 2D imaging is possible in
scanning mode, without increasing dose. Similarly, for tomographic images,
the approach allows single slice \ac{CT} or full 3D imaging in scanning mode.

Grating design and fabrication is nonstandard and involves a complex mask
design, as shown in Figure~\ref{Fig:grating_mask}. Each grating resides on a
silicon chip and has its specific structure length and curvature. For the
current experiments, a symmetric interferometer with a grating period of $p
= \SI{2.8}{\micro\metre}$ for all gratings has been used. The design energy
is $\SI{100}{\kilo\electronvolt}$ and the beam splitter grating periodically
shifts the phase by zero and $\pi$ at this energy~\cite{David2002}. Using
gold as the phase shifting material, a structure length of
$h_1 = \SI{19.8}{\micro \metre}$ is required. The analyzer grating is an absorption mask
for sensing slight changes of the interference pattern generated by the beam
splitter~\cite{Momose2003a}. With a structure length of $h_2 =
\SI{800}{\micro \metre}$
it can sufficiently attenuate X-rays up to energies of 
$\SI{160}{\kilo\electronvolt}$. Beam splitter and analyzer grating are
separated at the first fractional Talbot order~\cite{Weitkamp2005},
resulting in an intergrating distance of $\SI{158}{\milli\metre}$. The
source grating splits the relatively large focal spot ($\sim
\SI{1}{\milli\metre}$) into an array of individually coherent, but mutually
incoherent sources~\cite{Pfeiffer2006}. It is also made of gold structures
with a length of $h_0 = h_2 = \SI{800}{\micro \metre}$.
\begin{figure} [ht]
    \includegraphics[width=.6\textwidth]{figures/grating_mask.eps}
    \caption{Grating design mask for
        the edge-on illumination approach and \ac{SEM}
        image of the grating. The top part of the 4 inch wafer shows
        five grating chips. The
        gratings have different curvatures which are specific to the grating
        interferometer geometry. 
        The \ac{SEM} image shows the gold structures and the interrupting bridges
        that prevent the lamellae from collapsing~\cite{Kenntner2010}.}
        \label{Fig:grating_mask}
\end{figure}

@@

Edge-on illuminated grating interferometry breaks the current limitations of
phase-contrast imaging to the lower diagnostic energy range. Compact
geometries  using conventional X-ray sources for design energies well above
$\SI{100}{\kilo\electronvolt}$ could be realized with the approach. This
enables the examination of materials of higher density or thickness which
would be opaque at lower energies. Finally, the approach is not
limited to X-ray imaging, but could be easily applied to other grating-based
imaging modalities (e.g., with neutrons~\cite{Grunzweig2008}) where high
aspect ratios are required.

@@

\begin{thebibliography}{29}%

@@

@@

>>>>

X-ray radiography and \ac{CT} are standard imaging techniques in
materials and life sciences for the nondestructive examination of samples or
diagnostic tasks in medicine. The underlying contrast mechanism
relies on the different X-ray attenuation properties of different materials
or tissue types. The dominant physical effects contributing to attenuation
are the photoelectric effect and incoherent (Compton) scattering. Besides attenuation, the wave
nature of X-rays reveals another contrast mechanism, which is the phase
shift. The interaction contributing to phase shifts is coherent (Rayleigh)
scattering~\cite{Als-Nielsen2011}.

@@

Phase-sensitive
imaging is a desirable modality, as it delivers a complementary source of
contrast with respect to absorption by providing direct access 
to the electron density~\cite{Als-Nielsen2011}. Moreover, the combination with
attenuation enables the determination of the effective atomic
number~\cite{Qi2010}.
An enhanced
\ac{CNR} in images compared to attenuation for certain
materials or tissues~\cite{Pfeiffer2007a,McDonald2009} has also been
demonstrated.

The vast majority of phase-sensitive techniques, including crystal analyzer
based~\cite{Davis1995,Chapman1997} or interferometric~\cite{Bonse1965,Momose1996}
methods rely on X-ray beams of high spatial and temporal coherence, which is
available only at synchrotron sources. Inline
phase contrast~\cite{Snigirev1995,Wilkins1996,Cloetens1996} and
Talbot interferometry~\cite{Cloetens1997,David2002,Momose2003a} need high
spatial coherence but are available on polychromatic microfocus sources.
Phase-contrast imaging using X-ray beams of low temporal \emph{and}
spatial coherence such as conventional low-brilliance X-ray tubes have been
demonstrated with coded
apertures~\cite{Munro2012} and Talbot-Lau
interferometry~\cite{Pfeiffer2006}. Analyzer-based systems have been
recently extended to tube sources~\cite{Nesch2009,Parham2009}
but only at energies up
to the tungsten $K_\alpha$ line at~\SI{60}{\kilo\eV}.
In addition to phase sensitivity, analyzer-based, Talbot interferometry and
coded-aperture approaches also provide (with different retrieval mechanisms)
information about the integrated local small angle scattering power from
microscopic density fluctuations in a specimen~\cite{Pfeiffer2008}. This signal is known under the name of dark-field, scatter or visibility reduction contrast.

High-energy Talbot interferometry has been reported so far using a
synchrotron source at nominal energies of
$\SI{82}{\kilo\electronvolt}$~\cite{Willner2013} and 
$\SI{123}{\kilo\electronvolt}$~\cite{Ruiz2013}.
Using a low-brilliance X-ray tube, Talbot-Lau interferometry was
applied so far at $\SI{60}{\kilo\electronvolt}$ mean energy~\cite{Donath2009}.
Medical imaging applications may benefit from phase contrast at higher
energies: chest or abdominal
radiography or \ac{CT} require energies between $100$ and
$\SI{150}{\kilo\voltpeak}$. Other potential applications are homeland
security or chip failure analysis, which require high energies for the
visualization of materials of high density and atomic number.

Here we introduce a method for phase contrast imaging which works on
conventional X-ray sources, covers the entire diagnostic X-ray energy range
and is compatible with compact imaging arrangements. The approach is based
on
Talbot-Lau
interferometry~\cite{Pfeiffer2006} and employs an edge-on illumination approach for the grating design and
arrangement. 
Our solution removes one on the major hurdle which prevented grating interferometry to be applied at very high X-ray energies so far, namely the fabrication of gratings with high-aspect ratios.
 The aspect
ratio, given by
\begin{equation}
    \text{R} = \frac{2h}{p},
\end{equation}
where $p$ is the grating period and $h$ the  structure
height, is normally limited by the fabrication process, usually
photolithography~\cite{David2002} or X-ray lithography~\cite{Mohr2012} as grating
structures tend to collapse or deform (e.g. due to capillary forces) if
the aspect ratio is too high. For a given setup length these parameters
depend on the target energy $E$ according to $p \propto
1/\sqrt{E}$ and $h \propto E^3$, and therefore $\textnormal{R}
\propto E^{7/2}$~\cite{Momose2003a}. If at $E=\SI{25}{\kilo\electronvolt}$ an aspect ratio
for the absorption grating around $\textnormal{R}=30$ is necessary for a
reasonable length of the experimental arrangement, it would have to be at
least $128$ for $E=\SI{100}{\kilo\electronvolt}$. Moreover, when using a
broad spectrum, photons above the design energy should also be
efficiently blocked by the gratings, which requires even higher aspect ratios.
The largest aspect ratios achieved by current fabrication techniques~\cite{David2007,Kenntner2010} are around 60.

Our design introduces the edge-on
illuminated,  circularly aligned structures. Edge-on illumination
(figure~\ref{Fig:schematic}), as
opposed to face-on illumination, exploits the dimension along the grating
lines to form a high aspect ratio of the structures in the direction of the beam. The
effective structure height of the grating is then determined by the grating
dimension along the grating lines, which essentially allows arbitrarily high
aspect ratios. 
\begin{figure}[ht]
    \includegraphics[width=.7\textwidth]{figures/figure1.eps}
    \caption{Schematic of a grating
        interferometer for X-ray energies between 60 and
        \SI{150}{\kilo\electronvolt} in edge-on illumination mode. The
        aspect ratio is defined by the ratio of the travelling distance along the
        grating lines and the period and can be arbitrarily long. In order to avoid
        a reduction of the field of view, the grating structures are aligned on an
    arc.}
    \label{Fig:schematic}
\end{figure}

Increasing the aspect ratio of the gratings typically leads to a
reduction of the field of view due to the change of the grating transmission
function at high incident angles. In order to overcome this problem, the grating lines are circularly aligned 
with a radius equal to the distance to the source. This cannot be done with
the glancing angle approach~\cite{Stutman2012a}, and allows an arbitrarily
large field-of-view in a fan beam geometry.

The combination of edge-on illumination and circularly aligned structures
enables phase-contrast imaging at arbitrary design energies and with a
maximum field of view in the horizontal direction ($x$ direction). These
advantages come at the expense of a limited field of view in the vertical
direction ($y$ direction), which is, depending on the X-ray detector,
typically a few pixels. 
Radiographic 2D imaging can be obtained by scanning the sample or a thin fan
beam. The scanning technique has been demonstrated to deliver less dose than
the conventional approach based on the illumination of a large area. In
digital mammography, for instance, where dose is a critical issue, Philips'
MicroDose system combines a scanning approach with an highly collimated fan
beam~\cite{Aslund2007}. Thanks to the high collimation, the dose deposited
on  patients has been reported to be significantly lower than with other
instruments based on the illumination of a large area
detector~\cite{Oduko2010}.
Similarly, for tomographic images,
the approach allows single slice \ac{CT} or full 3D imaging in scanning mode.

Grating design and fabrication is nonstandard and involves a complex mask
design, as shown in Figure~\ref{Fig:grating_mask}. Multiple gratings can reside on a
silicon chip with their specific structure length and curvature. For the
current experiments, a symmetric interferometer with a grating period of $p
= \SI{2.8}{\micro\metre}$ for all gratings has been used. The design energy
is $\SI{100}{\kilo\electronvolt}$ and the beam splitter grating periodically
shifts the phase by zero and $\pi$ at this energy~\cite{David2002}. Using
gold as the phase shifting material, a structure length of
$h_1 = \SI{19.8}{\micro \metre}$ is required. The analyzer grating is an absorption mask
for sensing slight changes of the interference pattern generated by the beam
splitter~\cite{Momose2003a}. With a structure length of $h_2 =
\SI{800}{\micro \metre}$
it can sufficiently attenuate X-rays up to energies of 
$\SI{160}{\kilo\electronvolt}$. Beam splitter and analyzer grating are
separated at the first fractional Talbot order~\cite{Weitkamp2005},
resulting in an intergrating distance of $\SI{158}{\milli\metre}$. The
source grating splits the relatively large focal spot ($\sim
\SI{1}{\milli\metre}$) into an array of individually coherent, but mutually
incoherent sources~\cite{Pfeiffer2006}. It is also made of gold structures
with a length of $h_0 = h_2 = \SI{800}{\micro \metre}$.
\begin{figure} [ht]
    \includegraphics[width=.6\textwidth]{figures/grating_mask.eps}
    \caption{Grating design mask for
        the edge-on illumination approach and \ac{SEM}
        image of the grating. The top part of the 4 inch wafer shows
        five grating chips. The
        gratings have different curvatures which are specific to the grating
        interferometer geometry. 
        The \ac{SEM} image shows the gold structures and the interrupting bridges
        that prevent the lamellae from collapsing~\cite{Kenntner2010}.}
        \label{Fig:grating_mask}
\end{figure}

@@

Edge-on illuminated grating interferometry finally allows phase contrast
imaging to be performed at high energies ($>\SI{100}{\kilo\eV}$) on conventional X-ray
sources. With a circular distribution of the gratings, the diffracting and
absorbing structures can be matched to the divergent beam, even for very
compact geometries. Short, high-energy phase-contrast systems enable the
efficient investigation of high-density materials or large thickness,
adding information on electron density and integrated small angle scattering
power to the conventional absorption based signal. This will improve
material discrimination and density sensitivity capabilities in future
X-rays or even neutron~\cite{Grunzweig2008} investigations.

@@

\begin{thebibliography}{33}%

@@

\bibitem [{\citenamefont {Mohr}\ \emph {et~al.}(2012)\citenamefont {Mohr},
  \citenamefont {Grund}, \citenamefont {Kunka}, \citenamefont {Kenntner},
  \citenamefont {Leuthold}, \citenamefont {Meiser}, \citenamefont {Schulz},\
  and\ \citenamefont {Walter}}]{Mohr2012}%
  {Mohr}}, \bibinfo {author} {\bibfnamefont {T.}~\bibnamefont {Grund}},
  \bibinfo {author} {\bibfnamefont {D.}~\bibnamefont {Kunka}}, \bibinfo
  {author} {\bibfnamefont {J.}~\bibnamefont {Kenntner}}, \bibinfo {author}
  {\bibfnamefont {J.}~\bibnamefont {Leuthold}}, \bibinfo {author}
  {\bibfnamefont {J.}~\bibnamefont {Meiser}}, \bibinfo {author} {\bibfnamefont
  {J.}~\bibnamefont {Schulz}}, \ and\ \bibinfo {author} {\bibfnamefont
  {M.}~\bibnamefont {Walter}},\ }\href {\doibase 10.1063/1.4742267} {\ \textbf
  {\bibinfo {volume} {41}},\ \bibinfo {pages} {41} (\bibinfo {year}
  {2012})}\BibitemShut {NoStop}%

@@

\bibitem [{\citenamefont {Stutman}\ and\ \citenamefont
  {Finkenthal}(2012)}]{Stutman2012a}%
  {Stutman}}\ and\ \bibinfo {author} {\bibfnamefont {M.}~\bibnamefont
  {Finkenthal}},\ }\href {\doibase 10.1063/1.4748882} {\bibfield  {journal}
  {\bibinfo  {journal} {Applied physics letters}\ }\textbf {\bibinfo {volume}
  {101}},\ \bibinfo {pages} {91108} (\bibinfo {year} {2012})}\BibitemShut
\bibitem [{\citenamefont {{\AA slund, M}}(2007)}]{Aslund2007}%
  \bibfield  {author} {\bibinfo {author} {\bibnamefont {{\AA slund, M}}},\
  }\emph {\bibinfo {title} {{Digital Mammography with a Photon Counting
  Detector in a Scanned Multislit Geometry}}},\ \href@noop {} {Ph.D. thesis}
  (\bibinfo {year} {2007})\BibitemShut {NoStop}%
\bibitem [{\citenamefont {Oduko}\ \emph {et~al.}(2010)\citenamefont {Oduko},
  \citenamefont {Young},\ and\ \citenamefont {Burch}}]{Oduko2010}%
  \bibfield  {author} {\bibinfo {author} {\bibfnamefont {J.~M.}\ \bibnamefont
  {Oduko}}, \bibinfo {author} {\bibfnamefont {K.~C.}\ \bibnamefont {Young}}, \
  and\ \bibinfo {author} {\bibfnamefont {A.}~\bibnamefont {Burch}},\ }in\ \href
  {\doibase 10.1007/978-3-642-13666-5_49} {\emph {\bibinfo {booktitle} {Digital
  Mammography}}},\ \bibinfo {series} {Lecture Notes in Computer Science}, Vol.\
  \bibinfo {volume} {6136},\ \bibinfo {editor} {edited by\ \bibinfo {editor}
  {\bibfnamefont {J.}~\bibnamefont {Martí}}, \bibinfo {editor} {\bibfnamefont
  {A.}~\bibnamefont {Oliver}}, \bibinfo {editor} {\bibfnamefont
  {J.}~\bibnamefont {Freixenet}}, \ and\ \bibinfo {editor} {\bibfnamefont
  {R.}~\bibnamefont {Martí}}}\ (\bibinfo  {publisher} {Springer Berlin
  Heidelberg},\ \bibinfo {year} {2010})\ pp.\ \bibinfo {pages}
  {365--370}\BibitemShut {NoStop}%
